---
title: "ADUMonroe"
output: html_document
---


```{r}
library(tidycensus)
library(tidyverse)
library(readxl)
library(knitr)
library(survey)
library(srvyr)
library(ipumsr)
library(jtools)
library(weights)
library(fastDummies)
library(R.devices)
library(huxtable)
```

```{r}
ddi <- read_ipums_ddi("/Users/Amikula/Desktop/usa_00005.xml")
data <- read_ipums_micro(ddi) %>%
  filter(RENT > 0) %>%
  filter(FTOTINC > 0) %>%
    mutate(RACE = as_factor(RACE)) 
```

```{r}
group_by(data, YEAR)
group_data <- group_by(data, YEAR)
  
```

```{r}
data$White <- ifelse(data$RACE == 'White', 1, 0) 
data$Black <- ifelse(data$RACE == 'Black/African American/Negro', 1, 0)
data$Native <- ifelse(data$RACE == 'American Indian or Alaska Native', 1, 0)
data$Asian <- ifelse(data$RACE %in% c('Chinese', 'Japanese', 'Other Asian or Pacific Islander'), 1, 0)
data$Other <- ifelse(data$RACE == 'Other race, nec', 1, 0)
data$Multiple <- ifelse(data$RACE %in% c('Two major races', 'Three or more major races'), 1, 0)
```

```{r}
svy_data <- data %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = HHWT)
```

```{r}
corrs <- svycor(~FTOTINC + RENT + UNITSSTR + YEAR, design = svy_data, sig.stats = TRUE) 
```

```{r}
corrs$cors
```

```{r}
corrs$p.values
```

```{r}
corrs$std.err
```

```{r}
corrs_ci_low <- corrs$cors - 1.96*corrs$std.err
corrs_ci_upp <- corrs$cors + 1.96*corrs$std.err
```

```{r}
corrs_ci_low
```

```{r}
corrs_ci_upp
```

```{r}
corrs_low_df <- as_tibble(corrs_ci_low) %>%
  mutate(variable = c("Income", "Rent", "Units in building", "Year")) %>%
  rename(FTOTINC_low = FTOTINC,
         RENT_low = RENT,
         UNITSSTR_low = UNITSSTR,
         YEAR_low = YEAR)
corrs_upp_df <- as_tibble(corrs_ci_upp) %>%
  mutate(variable = c("Income", "Rent", "Units per structure", "Year")) %>%
  rename(FTOTINC_upp = FTOTINC,
         RENT_upp = RENT,
         UNITSSTR_upp = UNITSSTR,
         YEAR_upp = YEAR)
corrs_pretty_int <- left_join(corrs_low_df, corrs_upp_df) %>%
  mutate(`Income` = paste(prettyNum(FTOTINC_low, digits = 3), 
                                 " to ", 
                                 prettyNum(FTOTINC_upp, digits = 3))) %>%
  mutate(`Rent` = paste(prettyNum(RENT_low, digits = 3), 
                                 " to ", 
                                 prettyNum(RENT_upp, digits = 3))) %>%
  mutate(`Units in building` = paste(prettyNum(UNITSSTR_low, digits = 3), 
                                 " to ", 
                                 prettyNum(UNITSSTR_upp, digits = 3))) %>%
  mutate(`Year` = paste(prettyNum(YEAR_low, digits = 3), 
                                 " to ", 
                                 prettyNum(YEAR_upp, digits = 3))) %>%
  select(variable, `Income`, `Rent`, `Units in building`, `Year`)
  
kable(corrs_pretty_int)
```

```{r}
rent_model <- svyglm(RENT ~ YEAR, design = svy_data)
summary(rent_model)
```

The above regression predicts that the rent in Monroe County will rise by an average of $14.09 per year. It is statistically significant and strong in magnitude, although the line of best fit implies that, for much of the 19th century and earlier, the average monthly rent was a negative number, which is not an intuitive result.

```{r}
inc_model <- svyglm(RENT ~ FTOTINC, design = svy_data)
summary(inc_model)
```

The above regression implies that, for every dollar increase in family income, the rent the family pays decreases by a small amount. It is not a statistically significant relationship. Perhaps we could achieve a more usable result by expressing income in terms of thousands of dollars. 

```{r}
hu_model <- svyglm(RENT ~ UNITSSTR, design = svy_data)
summary(hu_model)
```

The above regression predicts that, for every 1-unit increase in the number of housing units in the structure, rents decline by about $7.66. While this is a significant magnitude, the relationship is not statistically significant. Notice that the intercept implies that structures with 1 unit have an average rent of (893.76 - 7.66 = ) 886.10 dollars per month.

```{r}
white_model <- svyglm(RENT ~ White, design = svy_data)
summary(white_model)
```

The above regression predicts that, all else equal, a white person has an average rent $64.05 higher than an average non-white person. It is a statistically significant relationship.

```{r}
black_model <- svyglm(RENT ~ Black, design = svy_data)
summary(black_model)
```

The above regression predicts that, all else equal, a black person has an average rent $148.72 lower than an average non-black person. It is a statistically significant relationship.

```{r}
native_model <- svyglm(RENT ~ Native, design = svy_data)
summary(native_model)
```

The above regression predicts that, all else equal, a Native American person has an average rent $273.82 lower than an average non-native person. It is a highly statistically significant relationship.

```{r}
asian_model <- svyglm(RENT ~ Asian, design = svy_data)
summary(asian_model)
```

The above regression predicts that, all else equal, an Asian person has an average rent $43.97 higher than an average non-Asian person. It is not a statistically significant relationship, likely because of a small sample size of Asians in Monroe County.

```{r}
other_model <- svyglm(RENT ~ Other, design = svy_data)
summary(other_model)
```

The above regression predicts that, all else equal, a person of races not otherwise included in the above analysis has an average rent $125.67 lower than an average "non-other" person. It is not a statistically significant relationship at the .05 level, likely because of a small sample size of people of "other races" in Monroe County. Given the relatively low p-value, however, we're somewhat confident that the relationship would be significant with a larger sample size.

```{r}
multiple_model <- svyglm(RENT ~ Multiple, design = svy_data)
summary(multiple_model)
```

The above regression predicts that, all else equal, a person of multiple races has an average rent $132.03 lower than an average person with one race. It is a statistically significant relationship.

Overall, the regressions with the race dummy variables probably reflect characteristics about the incomes of people of various races as well, not just factors like discrimination that are associated with the race variables alone. Once we get into multivariate analysis next week, it will be interesting to see how the magnitude and significance of these race effects change in our models.

```{r}
multivariate_model <- svyglm(RENT ~ YEAR + FTOTINC + UNITSSTR + Black + Native + Asian + Other + Multiple, design = svy_data)
summary(multivariate_model)
```

The above regression predicts that, all else equal, the rent will increase by an average of $14.69 per year. It is a highly statistically significant relationship. The income and units in structure variables continue to lack statistical significance. Relative to white people, blacks, Native Americans, and people of multiple races have statistically significantly lower rents. The rents of Asians and people of other races are not significantly different from those of whites, just as they weren't for the bivariate regression analysis. The signs on all of the variables remained the same compared to the bivariate regression analyses. Notably, however, the magnitude of the coefficient on the year variable is slightly larger than it was before, implying that some of the other variables are "working against" the yearly increases in rent, perhaps because the county is growing more racially diverse or having a larger concentration of units in large structures over time. 

```{r}
centered_data <- data %>%
  mutate(YEAR = YEAR - mean(YEAR, na.rm=TRUE),
         FTOTINC = FTOTINC - mean(FTOTINC, na.rm=TRUE),
         UNITSSTR = UNITSSTR - mean(FTOTINC, na.rm=TRUE))
```

```{r}
svy_centered <- centered_data %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = PERWT)
```

```{r}
centered_model <- svyglm(RENT ~ YEAR + FTOTINC + UNITSSTR + Black + Native + Asian + Other + Multiple, design = svy_centered)

export_summs(centered_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Centered"))
```

In the above model, we centered the data so that the mean value of each variable corresponds to a coefficient of zero and the intercept is more easily understandable as the expected value of rent given average values for all the other variables. As expected, this centered data model does not change the significance of any of our variables vis a vis the original multivariate model. It merely provides a more intuitive understanding of the relationship between the variables, allowing us to more easily calculate individual rent figures for a "typical" data point. The intercept tells us that, given average values for the other data points, the expected rent is about $899.05 per month. However, we're uncertain about how to interpret the dummy variables in the model given that they only have a value of either 0 or 1. Would the coefficients in this model reflect the share of the population of each racial category? If so, we're not sure how useful mean centering is for our data set.  

```{r}
data_scaled <- data %>%
  mutate(FTOTINC2 = FTOTINC/1000)
```

```{r}
svy_scaled <- data_scaled %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = PERWT)
```

```{r}
scaled_model <- svyglm(RENT ~ YEAR + FTOTINC2 + UNITSSTR + Black + Native + Asian + Other + Multiple, design = svy_scaled)

export_summs(scaled_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Scaled"))
```

In the above model, we scaled the income data so that the unit of change is one thousand dollars, not a single dollar. This makes our coefficient larger by a factor of 1,000, improving our ability to interpret the data in terms of a "practical" amount of money. However, the effect is still really small. Our model predicts that a 1,000-dollar increase in income reduces the household's rent by 1 cent. The scaling does not change the statistical significance. We're not sure this model is entirely helpful either - the income effect on rent is still small. Perhaps a non-linear transformation would fit the model better, as we explore below. 

```{r}
data_log <- data %>%
  mutate(logFTOTINC = log(FTOTINC, base = 2))
```

```{r}
svy_log <- data_log %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = PERWT)
```

```{r}
log_model <- svyglm(RENT ~ YEAR + logFTOTINC + UNITSSTR + Black + Native + Asian + Other + Multiple, design = svy_log)

export_summs(log_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Logged"))
```

In the above model, we transformed the income data using a log function with base 2, meaning that the unit of change is squaring income. This makes our coefficient very large and easily interpretable - squaring income is associated with a highly statistically significant $37.88 decrease in rent. This is surprising as for most goods, increasing income tends to increase spending, however, this is not the case for rental housing in Monroe County. Also, the Black variable is less statistically significant in this model than before, perhaps implying that there is some interaction going on between the Black and income variables. While the statistically significant result on the income variable is exciting, even more exciting is that this transformation apparently improves the fit of the model, the true indicator of a more appropriate model to use to understand the relationship between these variables. 

```{r}
interaction_model <- svyglm(RENT ~ YEAR + logFTOTINC + UNITSSTR + Black + Native + Asian + Other + Multiple + logFTOTINC*Black + logFTOTINC*Native + logFTOTINC*Asian + logFTOTINC*Other + logFTOTINC*Multiple, design = svy_log)

export_summs(interaction_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Interacted"))
```

In the above model, we interact each race dummy variable with the logged income variable. This allows us to see whether the relationship between income and median monthly rent is different for different races of people. We find that, relative to whites, both blacks and people of multiple races have a significantly lesser discount on their rent as their incomes increase, and for blacks the magnitude is positive overall. The results are not statistically significanat for Natives, Asians, and people of other races relative to whites. Also, the logFTOTINC variable by itself is still strongly statistically significant and negative in magnitude, demonstrating that, if you're white, your rent decreases on average as your income increases. We select this regression as our new preferred model, as the fit has improved compared to the regressions without the interaction terms, and because the model portrays some valuable information about the nuances of the relationship between income and rent payments. 
