---
title: "RMarkdownBloomington"
author: "Keana Flores"
date: "10/29/2021"
output: html_document
---
title: "Assignment One: Research Questions"
author: "Keana Flores, Lindsey Mayer, Andrew Mikula"
date: "10/25/2021"
output: html_document
---

*Research Question*

Did the adoption of an accessory dwelling unit (ADU) ordinance lower the cost of rent in Monroe County, Indiana? 


*Prior Research*

Crane (2020) found that ADUs “have increased the supply of relatively low-cost rental units in low-density neighborhoods” in California, but it is either illegal or undesirable for many homeowners to do so, thus limiting the potential for a significant price effect. Ramsey-Musolf (2018) is also skeptical of the ability of California’s ADUs to provide large-scale low-income housing, finding that, among the 759 ADUs constructed in California in the available dataset,  “no units were identified as available low-income housing.”  

*Data + Definition of Sample Population*

The sample population for this study is Monroe County, Indiana between 2009 and 2019. The analysis will include the following variables:
- Number of White individuals (ACS)
- Number of Black individuals (ACS)
- Number of Native American individuals (ACS)
- Number of Asian individuals (ACS)
- Number of Native Hawaiian or Pacific Islander individuals (ACS)
- Total population (ACS)
- Number of housing units (ACS)
- Median income (ACS)
- Monthly housing costs (ACS)
- Rental homes (ACS)

*References*

Rebecca Elisabeth Crane. 2020. “Is Granny in that Flat?: How Regulations Shape the Construction and Use of Accessory Dwelling Units in Los Angeles.” The California Digital Library. https://escholarship.org/uc/item/2wh204vz
 
Mike Hanks. April 4, 2019. “Bloomington opens door to more affordable housing units.” Adams Publishing Group. https://www.hometownsource.com/sun_current/community/bloomington/bloomington-opens-door-to-more-affordable-housing-units/article_f17b8092-5652-11e9-8951-47a5d871ac49.html 
 
Darrel Ramsey-Musolf. Accessory Dwelling Units as Low-Income Housing: California’s Faustian Bargain. Urban Sci. 2(3), 89. DOI: https://doi.org/10.3390/urbansci2030089
 
United States Census Bureau. American Community Survey, 5-year estimates. 2009-2019.

*Data*

```{r}
library(tidycensus)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(dplyr)
```

```{r}
ddi <- read_ipums_ddi("/Users/AMikula/Desktop/usa_00005.xml")
data <- read_ipums_micro(ddi) %>%
  na_if(9999999) %>%
  filter(RENT > 0) %>%
  mutate(RACE = as_factor(RACE))
```

```{r}
svy_data <- data %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = PERWT)
```

```{r}
units_mean <- svy_data %>%
  summarize(mean = survey_mean(UNITSSTR, vartype = c("ci", "var"), na.rm = TRUE))

income_mean <- svy_data %>%
  summarize(mean = survey_mean(FTOTINC, vartype = c("ci", "var"), na.rm = TRUE))

rent_mean <- svy_data %>%
  summarize(mean = survey_mean(RENT, vartype = c("ci", "var"), na.rm = TRUE))
```

```{r}
units_time_quants <- svy_data %>%
  summarize(quants = survey_quantile(UNITSSTR, 
                                     quantiles =  c(0.25, 0.5, 0.75),
                                     vartype = NULL)) 

income_quants <- svy_data %>%
  summarize(quants = survey_quantile(FTOTINC, 
                                     quantiles =  c(0.25, 0.5, 0.75),
                                     vartype = NULL)) 

rent_quants <- svy_data %>%
  summarize(quants = survey_quantile(RENT, 
                                     quantiles =  c(0.25, 0.5, 0.75),
                                     vartype = NULL))
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
