---
title: "ADUMonroe"
output: html_document
---


```{r}
library(tidycensus)
library(tidyverse)
library(readxl)
library(knitr)
library(survey)
library(srvyr)
library(ipumsr)
library(jtools)
library(weights)
library(fastDummies)
```

```{r}
ddi <- read_ipums_ddi("/Users/Amikula/Desktop/usa_00005.xml")
data <- read_ipums_micro(ddi) %>%
  filter(RENT > 0) %>%
    mutate(RACE = as_factor(RACE)) 
```

```{r}
group_by(data, YEAR)
group_data <- group_by(data, YEAR)
  
```

```{r}
data$White <- ifelse(data$RACE == 'White', 1, 0) 
data$Black <- ifelse(data$RACE == 'Black/African American/Negro', 1, 0)
data$Native <- ifelse(data$RACE == 'American Indian or Alaska Native', 1, 0)
data$Asian <- ifelse(data$RACE %in% c('Chinese', 'Japanese', 'Other Asian or Pacific Islander'), 1, 0)
data$Other <- ifelse(data$RACE == 'Other race, nec', 1, 0)
data$Multiple <- ifelse(data$RACE %in% c('Two major races', 'Three or more major races'), 1, 0)
```

```{r}
svy_data <- data %>%
  as_survey_design(ids = CLUSTER,
                   strata = STRATA,
                   weights = HHWT)
```

```{r}
corrs <- svycor(~FTOTINC + RENT + UNITSSTR + YEAR, design = svy_data, sig.stats = TRUE) 
```

```{r}
corrs$cors
```

```{r}
corrs$p.values
```

```{r}
corrs$std.err
```

```{r}
corrs_ci_low <- corrs$cors - 1.96*corrs$std.err
corrs_ci_upp <- corrs$cors + 1.96*corrs$std.err
```

```{r}
corrs_ci_low
```

```{r}
corrs_ci_upp
```

```{r}
corrs_low_df <- as_tibble(corrs_ci_low) %>%
  mutate(variable = c("Income", "Rent", "Units in building", "Year")) %>%
  rename(FTOTINC_low = FTOTINC,
         RENT_low = RENT,
         UNITSSTR_low = UNITSSTR,
         YEAR_low = YEAR)
corrs_upp_df <- as_tibble(corrs_ci_upp) %>%
  mutate(variable = c("Income", "Rent", "Units per structure", "Year")) %>%
  rename(FTOTINC_upp = FTOTINC,
         RENT_upp = RENT,
         UNITSSTR_upp = UNITSSTR,
         YEAR_upp = YEAR)
corrs_pretty_int <- left_join(corrs_low_df, corrs_upp_df) %>%
  mutate(`Income` = paste(prettyNum(FTOTINC_low, digits = 3), 
                                 " to ", 
                                 prettyNum(FTOTINC_upp, digits = 3))) %>%
  mutate(`Rent` = paste(prettyNum(RENT_low, digits = 3), 
                                 " to ", 
                                 prettyNum(RENT_upp, digits = 3))) %>%
  mutate(`Units in building` = paste(prettyNum(UNITSSTR_low, digits = 3), 
                                 " to ", 
                                 prettyNum(UNITSSTR_upp, digits = 3))) %>%
  mutate(`Year` = paste(prettyNum(YEAR_low, digits = 3), 
                                 " to ", 
                                 prettyNum(YEAR_upp, digits = 3))) %>%
  select(variable, `Income`, `Rent`, `Units in building`, `Year`)
  
kable(corrs_pretty_int)
```

```{r}
rent_model <- svyglm(RENT ~ YEAR, design = svy_data)
summary(rent_model)
```

The above regression predicts that the rent in Monroe County will rise by an average of $14.09 per year. It is statistically significant and strong in magnitude, although the line of best fit implies that, for much of the 19th century and earlier, the average monthly rent was a negative number, which is not an intuitive result.

```{r}
inc_model <- svyglm(RENT ~ FTOTINC, design = svy_data)
summary(inc_model)
```

The above regression implies that, for every dollar increase in family income, the rent the family pays decreases by a small amount. It is not a statistically significant relationship. Perhaps we could achieve a more usable result by expressing income in terms of thousands of dollars. 

```{r}
hu_model <- svyglm(RENT ~ UNITSSTR, design = svy_data)
summary(hu_model)
```

The above regression predicts that, for every 1-unit increase in the number of housing units in the structure, rents decline by about $7.66. While this is a significant magnitude, the relationship is not statistically significant. Notice that the intercept implies that structures with 1 unit have an average rent of (893.76 - 7.66 = ) 886.10 dollars per month.

```{r}
white_model <- svyglm(RENT ~ White, design = svy_data)
summary(white_model)
```

The above regression predicts that, all else equal, a white person has an average rent $64.05 higher than an average non-white person. It is a statistically significant relationship.

```{r}
black_model <- svyglm(RENT ~ Black, design = svy_data)
summary(black_model)
```

The above regression predicts that, all else equal, a black person has an average rent $148.72 lower than an average non-black person. It is a statistically significant relationship.

```{r}
native_model <- svyglm(RENT ~ Native, design = svy_data)
summary(native_model)
```

The above regression predicts that, all else equal, a Native American person has an average rent $273.82 lower than an average non-native person. It is a highly statistically significant relationship.

```{r}
asian_model <- svyglm(RENT ~ Asian, design = svy_data)d
summary(asian_model)
```

The above regression predicts that, all else equal, an Asian person has an average rent $43.97 higher than an average non-Asian person. It is not a statistically significant relationship, likely because of a small sample size of Asians in Monroe County.

```{r}
other_model <- svyglm(RENT ~ Other, design = svy_data)
summary(other_model)
```

The above regression predicts that, all else equal, a person of races not otherwise included in the above analysis has an average rent $125.67 lower than an average "non-other" person. It is not a statistically significant relationship at the .05 level, likely because of a small sample size of people of "other races" in Monroe County. Given the relatively low p-value, however, we're somewhat confident that the relationship would be significant with a larger sample size.

```{r}
multiple_model <- svyglm(RENT ~ Multiple, design = svy_data)
summary(multiple_model)
```

The above regression predicts that, all else equal, a person of multiple races has an average rent $132.03 lower than an average person with one race. It is a statistically significant relationship.

Overall, the regressions with the race dummy variables probably reflect characteristics about the incomes of people of various races as well, not just factors like discrimination that are associated with the race variables alone. Once we get into multivariate analysis next week, it will be interesting to see how the magnitude and significance of these race effects change in our models.

```{r}
multivariate_model <- svyglm(RENT ~ YEAR + FTOTINC + UNITSSTR + Black + Native + Asian + Other + Multiple, design = svy_data)
summary(multivariate_model)
```

The above regression predicts that, all else equal, the rent will increase by an average of $14.69 per year. It is a highly statistically significant relationship. The income and units in structure variables continue to lack statistical significance. Relative to white people, blacks, Native Americans, and people of multiple races have statistically significantly lower rents. The rents of Asians and people of other races are not significantly different from those of whites, just as they weren't for the bivariate regression analysis. The signs on all of the variables remained the same compared to the bivariate regression analyses. Notably, however, the magnitude of the coefficient on the year variable is slightly larger than it was before, implying that some of the other variables are "working against" the yearly increases in rent, perhaps because the county is growing more racially diverse or having a larger concentration of units in large structures over time. 
